---
title: "knnModel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Analyzing Breast Cancer Wisconsin (Diagnostic) Data Set using kNN

 http://archive.ics.uci.edu/ml

### load library 

```{r }
library(class)         # for knn()
library(gmodels)    # for CrossTable()
```

### load the data set

```{r }
wbcd <- read.csv("wdbc.csv", header=FALSE, stringsAsFactors = FALSE)

colnames(wbcd) <- c("id", "diagnosis", "radius_mean", "texture_mean", "perimeter_mean", "area_mean", "smoothness_mean","compactness_mean","concavity_mean", "concave_points_mean", "symmetry_mean", "fractal_dimension_mean", "radius_sd", "texture_sd", "perimeter_sd", "area_sd", "smoothness_sd", "compactness_s", "concavity_sd", "concave_points_sd", "symmetry_sd", "fractal_dimension_sd", "radius_worst", "texture_worst", "perimeter_worst", "area_worst", "smoothness_worst","compactness_worst","concavity_worst", "concave_points_worst", "symmetry_worst", "fractal_dimension_worst")
```

### view the layout of data
```{r }
str(wbcd)
```

### we have id in first column, this is not required to create a machine learning model. So we discard it
```{r }
wbcd <- wbcd[-1]
```

### change the lable of diagnosis column
```{r }
table(wbcd$diagnosis)
wbcd$diagnosis <- factor(wbcd$diagnosis, levels = c("B", "M"),
                        labels = c("Benign", "Malignant"))
```

### check the diagnosis proportion
```{r }
round(prop.table(table(wbcd$diagnosis)) * 100, digits = 1)
summary(wbcd[c("radius_mean", "area_mean", "smoothness_mean")])
```

### now we scale the value using min max normalization
```{r }
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

wbcd_n <- as.data.frame(lapply(wbcd[2:31], normalize))  #normalize is user defined
summary(wbcd_n$area_mean)
```

### create knn model, we choose k = 21 as sqrt(469)
```{r }
wbcd_train <- wbcd_n[1:469, ]
wbcd_test <- wbcd_n[470:569, ]
wbcd_train_labels <- wbcd[1:469, 1]
wbcd_test_labels <- wbcd[470:569, 1]
wbcd_test_pred <- knn(train = wbcd_train, test = wbcd_test,
                        cl = wbcd_train_labels, k = 21)
```

### evaluate the accuracy using CrossTable
```{r }
CrossTable(x = wbcd_test_labels, y = wbcd_test_pred,
           prop.chisq=FALSE)
```
